---
title: "Position analysis: evaluation, search, win probability"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Position analysis: evaluation, search, win probability}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/r-heller/multixoR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/r-heller/multixoR/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/r-heller/multixoR/actions/workflows/pkgdown.yaml/badge.svg)](https://r-heller.github.io/multixoR/)
[![CRAN status](https://www.r-pkg.org/badges/version/multixoR)](https://CRAN.R-project.org/package=multixoR)
[![Codecov test coverage](https://codecov.io/gh/r-heller/multixoR/branch/main/graph/badge.svg)](https://app.codecov.io/gh/r-heller/multixoR?branch=main)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/multixoR)](https://cran.r-project.org/package=multixoR)
[![CRAN downloads total](https://cranlogs.r-pkg.org/badges/grand-total/multixoR)](https://cran.r-project.org/package=multixoR)
[![License: MIT](https://img.shields.io/badge/license-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 6, fig.height = 4)
library(multixoR)
```

## One evaluator, many consumers

The single source of truth for "how good is this position?" is
`mxo_evaluate()`. It counts existence-gated `m`-lines (length-`k` extents
with `m` marks of one colour and the rest empty) per axis class and sums
a weighted contribution from each side's perspective.

```{r evaluate}
g <- mxo_example_game()
mxo_evaluate(g)                    # X perspective by default
mxo_evaluate(g, player = 2L)       # O perspective is the negation
```

The raw feature vector — same input the heuristic and the calibrator
share — is available as an internal helper:

```{r features}
head(multixoR:::.mxo_line_features(g), 8L)
```

## Searcher and MCTS

`mxo_search()` implements negamax + α-β with a configurable
`branch_policy` (`all` / `promising` / `none`). The branching factor of
the multiverse is large, so `none` is the practical default for fast
local lookahead.

```{r search}
small <- mxo_new_game(n = 3L, k = 3L)
small <- mxo_move(small, "present", 0L, 0L, 0L)
res <- mxo_search(small, depth = 1L, branch_policy = "none")
res$move
```

`mxo_mcts()` is the real-time path: a light UCT with random or
heuristic rollouts. It's seedable for reproducibility.

```{r mcts}
set.seed(1L)
mc <- mxo_mcts(small, iterations = 30L,
               rollout = "random", branch_policy = "none")
mc$move
```

## Win probability, calibrated

`mxo_win_prob()` maps either the evaluator score (via the package's
internal calibrator, see `data-raw/make_calibrator.R`) or the MCTS visit
distribution to a `[0, 1]` win probability for a chosen player.

```{r win-prob}
mxo_win_prob(small, player = 1L, method = "heuristic")
```

## Move rating

`mxo_rate_moves()` is the type-stable per-move tibble the visualisation
stack and the Shiny app consume. The `label` column applies
chess-engine-style thresholds on the win-prob drop vs the best move.

```{r rate}
head(mxo_rate_moves(small, method = "heuristic"), 6L)
```

`mxo_ai_move()` is the convenience wrapper for the app's AI player; the
`difficulty` knob maps to search depth and MCTS iteration counts.

```{r ai-move}
mxo_ai_move(small, difficulty = "easy", seed = 1L)
```
