---
title: "Getting started with multixoR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with multixoR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 6, fig.height = 4)
```

`multixoR` is a five-dimensional, multiverse variant of tic-tac-toe. Two
players place marks on a 4×4×4 spatial cube; play also extends across a
**time** axis (successive board states) and a **timeline** axis (parallel,
branching universes). A player wins by making a length-3 line of their own
colour along any axis or diagonal — including lines that traverse time or
timelines.

```{r quickstart}
library(multixoR)

g <- mxo_new_game()
mxo_to_move(g)
nrow(mxo_legal_moves(g))
```

## Two kinds of move

A **present** move places the player's mark on an empty cell of the
current present board and advances time by one step.

```{r present}
g <- mxo_move(g, "present", L_src = 0L, t_src = 0L, idx = 0L)   # X at (0,0,0)
g <- mxo_move(g, "present", L_src = 0L, t_src = 1L, idx = 16L)  # O somewhere
mxo_history(g)
```

A **branch** move places into an empty cell of a *past* board and spawns
a new timeline that copies the source board plus the new mark. The source
timeline is left untouched.

```{r branch}
g <- mxo_move(g, "branch", L_src = 0L, t_src = 0L, idx = 63L)
mxo_timelines(g)
```

## Analysis

Every analysis function in the package consumes the engine's state object
directly — there is no parallel implementation.

```{r analysis}
mxo_evaluate(g)
mxo_win_prob(g, method = "heuristic")
```

`mxo_rate_moves()` returns the type-stable per-move table the visualisation
stack and the Shiny app consume; for the 4³ default it can be slow under
the pure-R engine, so this vignette demonstrates it on a 3³ position.

```{r rating}
small <- mxo_new_game(n = 3L, k = 3L)
small <- mxo_move(small, "present", 0L, 0L, 0L)
head(mxo_rate_moves(small, method = "heuristic"), 6L)
```

## Visualisation

`autoplot()` is the ergonomic dispatcher; the more specific renderers are
also exported.

```{r viz}
library(ggplot2)
autoplot(g)                                # multiverse overview
autoplot(g, type = "board", L = 0L, t = 0L)
autoplot(g, type = "threats")
```

## App

`mxo_run_app()` launches a bundled Shiny app for interactive play and
analysis. Requires the optional packages `shiny`, `bslib`, and `DT` (in
`Suggests`).

```{r app, eval = FALSE}
mxo_run_app()
```
