---
title: "Rules and 5D geometry"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rules and 5D geometry}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 6, fig.height = 4)
library(multixoR)
library(ggplot2)
```

## Five axes

`multixoR` is played in five dimensions. Three are spatial (`x`, `y`, `z`)
and define a `4×4×4` cube of 64 cells. The fourth axis is **time** (`t`),
a sequence of board snapshots within one timeline. The fifth is the
**timeline** (`L`) axis: every branching move spawns a new universe whose
boards live at parallel `L` labels.

A canonical address is the 5-tuple `(L, t, x, y, z)`. The engine works
generically over `(n, d_spatial, k)` — the default `4³` / `k = 3` is the
declared configuration, never a hardcoded constant.

```{r geometry}
cfg <- mxo_config(mxo_new_game())
cfg
```

## Directions and winning lines

A direction is a 5-vector in `{-1, 0, +1}` minus the all-zero vector,
canonicalised so the first non-zero component is positive. The total
number of canonical directions in `d_spatial + 2 = 5` axes is
`(3^5 − 1) / 2 = 121`.

```{r dirs}
nrow(multixoR:::.mxo_directions(3L))
```

A winning line is any three same-colour, existence-gated cells along a
canonical direction. The §5.2.1 amendment in the spec restricts wins to
extents that **pass through the most recent placement** — propagated
cells alone never declare a win.

## Branching across timelines

```{r branch}
g <- mxo_new_game()
g <- mxo_move(g, "present", 0L, 0L, 0L)   # X at (0,0,0)
g <- mxo_move(g, "present", 0L, 1L, 16L)  # O somewhere
g <- mxo_move(g, "branch",  0L, 0L, 63L)  # X branches to L1
mxo_timelines(g)
```

`mxo_plot_multiverse()` lays every board out on a timeline × time grid
and draws the branch connectors:

```{r multiverse-plot}
mxo_plot_multiverse(g)
```

## Three timeline-axis placements → win

A pure `dL`-axis win requires three placements at the **same** `(t, idx)`
on three adjacent timelines. The example below branches three times from
the empty initial board at the same spatial cell.

```{r timeline-win}
w <- mxo_new_game()
w <- mxo_move(w, "present", 0L, 0L, 0L)
w <- mxo_move(w, "present", 0L, 1L, 1L)
w <- mxo_move(w, "branch",  0L, 0L, 63L)
w <- mxo_move(w, "present", 0L, 2L, 16L)
w <- mxo_move(w, "branch",  0L, 0L, 63L)
w <- mxo_move(w, "present", 0L, 3L, 17L)
w <- mxo_move(w, "branch",  0L, 0L, 63L)
mxo_status(w)$status
```

The same engine, geometry, and win-check produce all five axis-class
wins; `mxo_plot_threats()` shows the per-axis-class threatened-line count
that drives strategic play.

```{r threats}
mxo_plot_threats(g, player = 1L, min_marks = 1L)
```
