---
title: "Self-play, simulation, and the §12 timeline-win finding"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Self-play, simulation, and the §12 timeline-win finding}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>",
                      fig.width = 6, fig.height = 4)
library(multixoR)
```

## Policies

A policy is a parameterised picker function over Stack B's searchers.

```{r policies}
mxo_policy("random")
mxo_policy("heuristic", branch_policy = "none")
```

## Single self-play game

`mxo_self_play()` returns an `mxo_game_record` that includes the per-ply
heuristic score and the heuristic win probability (X perspective). Those
two columns are exactly what the win-probability curve consumes.

```{r self-play}
rec <- mxo_self_play(
  mxo_policy("random"), mxo_policy("random"),
  config = mxo_config_default(n = 3L, k = 3L, ply_cap = 8L),
  seed = 1L
)
rec
head(as_tibble.mxo_game_record(rec), 6L)
```

```{r winprob-curve}
mxo_plot_win_prob(rec)
```

## Batch simulation

`mxo_simulate()` runs many self-play games with reproducible per-game
sub-seeds. The summary surfaces the §12 stress-test number — the
**fraction of decisive games whose winning line crosses timelines** —
the diagnostic the orchestrator flagged as the rules' main balance
concern.

```{r batch}
sim <- mxo_simulate(
  mxo_policy("random", branch_policy = "all"),
  mxo_policy("random", branch_policy = "all"),
  n_games = 20L,
  config = mxo_config_default(n = 3L, k = 3L, ply_cap = 8L),
  seed = 7L, record_eval = FALSE, progress = FALSE
)
summary(sim)
```

`mxo_timeline_win_rate()` returns the same number with a per-axis-class
breakdown.

```{r tlrate}
mxo_timeline_win_rate(
  mxo_policy("random"), mxo_policy("random"),
  n_games = 20L,
  config = mxo_config_default(n = 3L, k = 3L, ply_cap = 8L),
  seed = 7L
)
```

## Openings heatmap

`mxo_opening_table()` plays X's first move from each spatial cell against
a fixed opponent. `mxo_plot_opening()` renders the resulting win-rate
heatmap as Z-slices.

```{r opening}
tab <- mxo_opening_table(
  opponent = mxo_policy("random"),
  n_games_per_cell = 2L,
  config = mxo_config_default(n = 3L, k = 3L, ply_cap = 6L),
  seed = 1L
)
mxo_plot_opening(tab, n = 3L, d_spatial = 3L)
```

## Calibration

`mxo_make_calibration_data()` + `mxo_fit_calibration()` produce the
mapping the engine's win-probability uses internally. The script that
fits the package default lives at `data-raw/make_calibrator.R` and ships
the result to `R/sysdata.rda`.
