---
title: "2. Playing your first game"
description: >
  Turn order, present moves, the cell index in action, reading the board after
  every ply, and the canonical move notation.
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 7, fig.height = 4.5, dpi = 120, out.width = "100%"
)
library(multixoR)
library(ggplot2)
```

In [part 1](tutorial-1-the-board.html) we met the board. Now we actually play.
This page covers ordinary turn-by-turn play within a single timeline -- the
**present** move -- and leaves branching for [part 3](tutorial-3-branching.html).

## Starting a game

```{r new}
g <- mxo_new_game()
mxo_to_move(g)            # 1 = X moves first
nrow(mxo_legal_moves(g))  # every empty cell of the present board is legal
```

Player `1` is **X** and moves first; player `2` is **O**. On an empty cube all
64 cells are available.

## The present move

A *present* move places your mark on an empty cell of the current present
board and advances time by one step. The call is

```r
mxo_move(game, "present", L_src, t_src, idx)
```

where `(L_src, t_src)` is the board you are moving on -- `L = 0`, the current
time `t`, for ordinary play -- and `idx` is the cell. Let's make the first move:
X takes the corner `idx = 0`, i.e. `(0,0,0)`.

```{r move1}
g <- mxo_move(g, "present", L_src = 0L, t_src = 0L, idx = 0L)   # X (0,0,0)
mxo_plot_board(g)         # defaults to L = 0 and the present board
```

`mxo_plot_board()` with no `t` shows the present snapshot, so X's mark is
already there. It is now O's turn:

```{r move2}
mxo_to_move(g)
g <- mxo_move(g, "present", 0L, 1L, 16L)  # O (0,0,1): one step along z
mxo_plot_board(g)
```

Note the `t_src` climbed from `0` to `1`: each present move advances time, so
you always play on the latest board. X replies in the same `x`-row, building
toward a line:

```{r move3}
g <- mxo_move(g, "present", 0L, 2L, 1L)   # X (1,0,0)
mxo_plot_board(g)
```

## Reading the history

Every ply is recorded. `mxo_history()` returns the full move table:

```{r history}
mxo_history(g)
```

`mxo_status()` summarises where the game stands:

```{r status}
mxo_status(g)
```

## Move notation

multixoR has a compact text notation for plies (rules section 8). Serialize
any record row with `mxo_format_ply()`:

```{r notation}
h <- mxo_history(g)
for (i in seq_len(nrow(h))) cat(mxo_format_ply(as.list(h[i, ])), "\n")
```

The form is `PLAYER kind @ (L,t) [idx]`. `mxo_parse_ply()` is the exact inverse,
turning a string back into a record:

```{r parse}
mxo_parse_ply("X present @ (0,2) [1]")
```

## Legal moves at any point

`mxo_legal_moves()` always lists exactly the moves available from the current
state -- the engine never lets you play an illegal one (an occupied cell, for
instance, is rejected):

```{r legal}
head(mxo_legal_moves(g), 5)
```

So far we have only played in the present, in a single timeline. The next page
introduces the move that makes multixoR a multiverse game.

---

**Previous:** [1. The board and the five axes](tutorial-1-the-board.html) &nbsp;|&nbsp;
**Next:** [3. Branching into the past →](tutorial-3-branching.html)
