---
title: "3. Branching into the past"
description: >
  The branch move: placing into a past board to spawn a new parallel timeline,
  and visualising the resulting multiverse with the overview and tree plots.
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 7, fig.height = 4.5, dpi = 120, out.width = "100%"
)
library(multixoR)
library(ggplot2)
```

So far ([part 2](tutorial-2-first-game.html)) we only played present moves in a
single universe. The **branch** move is what turns multixoR into a multiverse
game.

## What a branch does

A branch move places your mark into an empty cell of a **past** board. Instead
of editing history, it spawns a brand-new, parallel **timeline** that copies
that past board *plus* your new mark. The original timeline is left completely
untouched. Branching is unrestricted -- you may branch from any reachable past
board, as often as you like.

The call mirrors a present move, but `kind = "branch"` and `(L_src, t_src)`
names the *past* board you are branching from:

```r
mxo_move(game, "branch", L_src, t_src, idx)
```

## A worked branch

Let's play three present moves and then branch from a past board. We start in
timeline `L = 0`:

```{r build}
g <- mxo_new_game()
g <- mxo_move(g, "present", 0L, 0L, 0L)   # X (0,0,0)
g <- mxo_move(g, "present", 0L, 1L, 5L)   # O (1,1,0)
g <- mxo_move(g, "present", 0L, 2L, 1L)   # X (1,0,0)
mxo_timelines(g)
```

There is one timeline so far. Now X -- rather, O is to move -- branches from the
*past* board `(L = 0, t = 1)` by placing at the far corner `idx = 63`:

```{r branch}
g <- mxo_move(g, "branch", L_src = 0L, t_src = 1L, idx = 63L)
mxo_timelines(g)
```

A second timeline, `L = 1`, has appeared. It contains the board as it stood at
`t = 1` in timeline 0, plus O's new mark at the corner -- while timeline 0 still
carries on with its own history.

## Seeing the multiverse

`mxo_plot_multiverse()` lays every board out on a **timeline x time** grid and
draws the branch connectors, so you can see at a glance where universes split:

```{r multiverse}
mxo_plot_multiverse(g)
```

`mxo_plot_tree()` shows the same branching as a tree of board states, which is
often the clearest way to follow how the timelines relate:

```{r tree}
mxo_plot_tree(g)
```

## Why branching matters

Branching is not a gimmick: a winning line may run **across** timelines (the
`L` axis) as readily as through space or time. That means a branch can set up a
threat that simply does not exist in ordinary tic-tac-toe, and a careless
opponent can be beaten in a universe they were not even watching. We make a
timeline-axis win concrete in the next page.

---

**Previous:** [2. Playing your first game](tutorial-2-first-game.html) &nbsp;|&nbsp;
**Next:** [4. Winning across space, time and timelines →](tutorial-4-winning.html)
