---
title: "4. Winning across space, time and timelines"
description: >
  The win condition, the five classes of winning line, a worked spatial win and
  a worked timeline win, the placement-anchored rule, and reading threats.
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 7, fig.height = 4.5, dpi = 120, out.width = "100%"
)
library(multixoR)
library(ggplot2)
```

You now know how to play present moves ([part 2](tutorial-2-first-game.html))
and how to branch ([part 3](tutorial-3-branching.html)). This page is the heart
of the game: **how you win.**

## The win condition

A player wins by completing a straight line of `k = 3` of their own marks along
**any** of the game's axes or diagonals. A "direction" is a 5-vector with
entries in `{-1, 0, +1}` (excluding all-zero), canonicalised so the first
non-zero component is positive. In five axes there are

$$\frac{3^5 - 1}{2} = 121$$

such canonical directions:

```{r dirs}
nrow(multixoR:::.mxo_directions(3L))
```

Because a line can mix the spatial, time, and timeline axes, multixoR sorts
winning (and threatening) lines into four **axis classes**: `spatial`, `time`,
`timeline`, and `mixed`.

## The placement-anchored rule

One subtlety is essential (rules section 5.2.1). When you place a mark, it
exists on that board *and propagates forward* to later boards in the same
timeline. But a win must pass through your **most recent placement** --
propagated cells alone never declare a win. This is why you cannot simply let a
mark "ride forward" through time to complete a line for free: the line has to be
anchored by the move you just made.

## A spatial win

The simplest win is three in a row along a single spatial axis on one board.
Here X builds the `x`-row `(0,0,0)-(1,0,0)-(2,0,0)` while O plays elsewhere:

```{r spatial}
s <- mxo_new_game()
s <- mxo_move(s, "present", 0L, 0L, 0L)    # X (0,0,0)
s <- mxo_move(s, "present", 0L, 1L, 16L)   # O elsewhere
s <- mxo_move(s, "present", 0L, 2L, 1L)    # X (1,0,0)
s <- mxo_move(s, "present", 0L, 3L, 17L)   # O elsewhere
s <- mxo_move(s, "present", 0L, 4L, 2L)    # X (2,0,0) -> completes the line
mxo_status(s)
```

The status reports `x_win` for player 1. The winning line was completed on the
board at `t = 4` (the moment of the final placement), and `mxo_plot_board()`
highlights that line for a finished game:

```{r spatial-plot}
mxo_plot_board(s, L = 0L, t = 4L)
```

## A timeline win

Now the move that has no analogue in ordinary tic-tac-toe: a win along the
**timeline** axis. It requires three placements at the *same* `(t, idx)` on
three adjacent timelines. We branch from the empty opening board at the same
corner cell three times:

```{r timeline}
w <- mxo_new_game()
w <- mxo_move(w, "present", 0L, 0L, 0L)
w <- mxo_move(w, "present", 0L, 1L, 1L)
w <- mxo_move(w, "branch",  0L, 0L, 63L)   # X into L1 corner
w <- mxo_move(w, "present", 0L, 2L, 16L)
w <- mxo_move(w, "branch",  0L, 0L, 63L)   # X into L2 corner
w <- mxo_move(w, "present", 0L, 3L, 17L)
w <- mxo_move(w, "branch",  0L, 0L, 63L)   # X into L3 corner -> timeline win
mxo_status(w)$status
```

The three corner marks sit at the same spatial cell on timelines 1, 2 and 3 --
a line straight along the `L` axis. The multiverse view shows them stacked
across parallel universes:

```{r timeline-plot}
mxo_plot_multiverse(w)
```

## The remaining classes: time and mixed

`time`-axis and `mixed` lines exist too -- a `time` line runs through the same
cell across successive boards, and a `mixed` line combines spatial steps with
time and/or timeline steps. Because of the placement-anchored rule above,
constructing these by hand takes care, but the engine detects all five classes
with the same machinery. The practical way to see them coming is to watch your
**threatened lines**.

## Reading threats

`mxo_plot_threats()` counts, per axis class, how many lines a player has
partially completed -- the lines that could become wins. Watching this for both
players is the core of multixoR strategy:

```{r threats}
g <- mxo_example_game()
mxo_plot_threats(g, player = 1L, min_marks = 1L)
```

We turn those threats into actual strategy -- and hand the job to the built-in
AI -- in the next page.

---

**Previous:** [3. Branching into the past](tutorial-3-branching.html) &nbsp;|&nbsp;
**Next:** [5. Strategy, evaluation and AI →](tutorial-5-strategy-ai.html)
