---
title: "5. Strategy, evaluation and AI"
description: >
  Reading positions with the heuristic evaluator and win probability, rating
  every legal move, and the search and MCTS engines behind the built-in AI.
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 7, fig.height = 4.5, dpi = 120, out.width = "100%"
)
library(multixoR)
library(ggplot2)
```

[Part 4](tutorial-4-winning.html) showed how lines complete. This page is about
*judging* a position before it is won, and letting the package play for you.
Every analysis function here consumes the engine's state object directly --
there is one engine and one evaluator behind the app, the simulations, and the
AI.

## Evaluating a position

`mxo_evaluate()` returns a single heuristic score from the side-to-move's point
of view. It weighs each player's threatened lines across all axis classes:

```{r evaluate}
g <- mxo_example_game()
mxo_evaluate(g)
```

`mxo_win_prob()` maps that score onto a calibrated `[0, 1]` win probability,
which is easier to reason about than a raw heuristic:

```{r winprob}
mxo_win_prob(g, method = "heuristic")
```

## Rating every legal move

`mxo_rate_moves()` is the type-stable per-move table the visualisation stack
and the Shiny app consume: it scores and ranks every legal move. On the full
`4^3` cube this is heavy under the pure-R engine, so -- as in the package's own
vignettes -- we demonstrate it on a smaller `3^3` position:

```{r rate}
small <- mxo_new_game(n = 3L, k = 3L)
small <- mxo_move(small, "present", 0L, 0L, 0L)
head(mxo_rate_moves(small, method = "heuristic"), 6)
```

The `rank` and `label` columns make it easy to surface the best moves -- exactly
what the app's overlays do.

## Searching ahead

`mxo_search()` runs alpha-beta negamax to a fixed depth. The `branch_policy`
argument controls whether the search is allowed to consider branch moves
(`"none"`, `"promising"`, or `"all"`), which matters a great deal because
branching explodes the move count:

```{r search}
mxo_search(g, depth = 1L, branch_policy = "none")$move
```

For deeper, anytime search there is `mxo_mcts()`, a Monte-Carlo tree search.
It is randomised, so we pass a `seed` for reproducibility and keep the example
small and fast (a `3^3` board, a modest iteration budget, random rollouts):

```{r mcts}
mcts <- mxo_mcts(
  mxo_new_game(n = 3L, k = 3L),
  iterations = 40L, rollout = "random", seed = 1L
)
mcts
```

## Just give me a move

For casual play you do not need any of the above directly. `mxo_ai_move()`
wraps the search behind three difficulty levels and returns a single move; pass
a `seed` to make it reproducible:

```{r ai}
mxo_ai_move(g, difficulty = "easy", seed = 1L)
mxo_ai_move(g, difficulty = "medium", seed = 1L)
```

## Strategy in a nutshell

- **Count threats, not marks.** A position is strong when you have more
  near-complete lines than your opponent -- `mxo_plot_threats()` and
  `mxo_evaluate()` both encode this.
- **Branches are weapons.** A timeline-axis or mixed threat can come from a
  universe your opponent is ignoring. Watch *all* timelines.
- **Respect the placement rule.** Because a win must pass through your latest
  placement, tempo matters: you cannot bank a propagated line for later.

Everything on this page is also available interactively -- which is the subject
of the final page.

---

**Previous:** [4. Winning across space, time and timelines](tutorial-4-winning.html) &nbsp;|&nbsp;
**Next:** [6. Playing in the Shiny app →](tutorial-6-the-app.html)
