---
title: "1. The board and the five axes"
description: >
  Where multixoR is played: the 4x4x4 spatial cube, the cell index, and the
  two extra axes - time and timeline - that make the game five-dimensional.
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, comment = "#>",
  fig.width = 7, fig.height = 4.5, dpi = 120, out.width = "100%"
)
library(multixoR)
library(ggplot2)
```

This is the first page of a six-part tutorial on how to play **multixoR**, a
five-dimensional, multiverse variant of tic-tac-toe. We start with the
question every player asks first: *where, exactly, am I placing my marks?*

```{r load, eval = FALSE}
install.packages("remotes")
remotes::install_github("r-heller/multixoR")
library(multixoR)
```

## The spatial cube

A single multixoR board is not a 3x3 grid. It is a `4x4x4` **cube** of 64
cells. Three of the game's five axes are these spatial axes -- `x`, `y`, and
`z`.

Every game carries its geometry with it. The default configuration is the
canonical `4^3` cube with a winning-line length of `k = 3`:

```{r config}
g <- mxo_new_game()
mxo_config(g)
```

The engine is generic over `(n, d_spatial, k)` -- nothing in the rules hard-codes
the number 4 -- but the canonical cube is the default, and the one this
tutorial uses.

### The cell index

Rather than typing three coordinates for every move, multixoR addresses each
cell of the cube with a single integer `idx` in `0..63`. The mapping is

$$\texttt{idx} = x + n\,y + n^2 z \qquad (n = 4),$$

so `x` runs fastest. A few examples:

```{r idx-table}
n <- 4L
cells <- expand.grid(x = 0:(n - 1), y = 0:(n - 1), z = 0:(n - 1))
cells$idx <- with(cells, x + n * y + n^2 * z)
head(cells[order(cells$idx), c("idx", "x", "y", "z")], 8)
```

So `idx = 0` is the corner `(0,0,0)`, `idx = 1` is `(1,0,0)`, `idx = 16` is
`(0,0,1)` -- one step along `z` -- and `idx = 63` is the far corner `(3,3,3)`.

### Seeing the cube

`mxo_plot_board()` draws one board. The default `"slices"` view unfolds the
cube into its four `z`-layers, which is the easiest way to read a position:

```{r board-slices}
mxo_plot_board(g, L = 0L, t = 0L)
```

The `"cube"` view renders the same board as an interactive 3-D plotly object --
drag to rotate it:

```{r board-cube}
mxo_plot_board(g, L = 0L, t = 0L, view = "cube")
```

An empty opening board is not very interesting. Here is the bundled example
game's opening cube, which already has a handful of marks:

```{r board-cube-example}
mxo_plot_board(mxo_example_game(), L = 0L, t = 0L, view = "cube")
```

## The fourth axis: time

multixoR is played in turns, and each board is a *snapshot* at one moment.
The sequence of snapshots within a single universe is the **time** axis, `t`.
Placing a mark produces a new board at the next time step; earlier snapshots
are never overwritten. We explore time properly in
[part 2](tutorial-2-first-game.html).

## The fifth axis: timeline

The last axis is what makes multixoR a *multiverse* game. Placing a mark into
an empty cell of a **past** board spawns a brand-new, parallel universe -- a new
**timeline**, labelled `L` -- that copies the past board plus your new mark,
leaving the original timeline untouched. Branching is covered in
[part 3](tutorial-3-branching.html).

## The full address

Putting it together, every cell in a multixoR game has a five-part address:

$$(L,\; t,\; x,\; y,\; z) \;=\; (\text{timeline},\; \text{time},\; \text{idx}).$$

Three spatial coordinates collapse into one `idx`, and the two extra axes `L`
and `t` say *which board* that cell lives on. Those five numbers are the five
dimensions in "five-dimensional tic-tac-toe", and a winning line may run along
any of them.

---

**Next:** [2. Playing your first game →](tutorial-2-first-game.html)
